---
title: "Data Analysis - Lab-3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In this homework, you will do some data analysis using R for the Forest Fire Data described
http://archive.ics.uci.edu/ml/machine-learningdatabases/forest-fires/forestfires.csv

The dataset is used to find the relationship between the burned area of forest fires and meterological data.

1. Import the data into R.

```{r forestfires}
ff.df <- read.csv("C:/Development/UCSC/Data Analysis/forestfires.csv", header = TRUE)
head(ff.df)
```

a. How many observations are there in the dataset?

```{r q1-a}
nrow(ff.df)
```

b. How many observations are there with a fire? 
(i.e.,	area > 0)

```{r q1-b}
x.sub <- subset(ff.df, area > 0)
nrow(x.sub)
#head(x.sub)
```

c. How many	observations are there with a rain?	
(i.e., rain > 0)

```{r q1-c}
x.sub <- subset(ff.df, rain > 0)
nrow(x.sub)
x.sub
```

d. How many	observations are there with both a fire and a	rain?	

```{r q1-d}
x.sub <- subset(ff.df, area > 0 & rain > 0)
nrow(x.sub)
#x.sub
```

2. Show	the columns	month, day, area of the all	the	observations.

```{r q2}
x.sub <- subset(ff.df, select = c("month", "day", "area"))
#nrow(x.sub)
x.sub
```

3. Show	the columns month, day, area of the observations with	a	fire.

```{r q3}
x.sub <- subset(ff.df, area > 0, select = c("month", "day", "area"))
#nrow(x.sub)
x.sub
```

4. How large are the five largest fires? (i.e., having largest area)

```{r q4}
x.sub <- ff.df[order(-ff.df$area),]
x.sub[1:5,]
```

a. What	are	the	corresponding	month, temp, RH, wind, rain, and area?

```{r q4-a}
x.sub <- ff.df[order(-ff.df$area),]
x.sub <- subset(x.sub, select = c("month", "temp", "RH", "wind", "rain", "area"))
x.sub[1:5,]
```

b. Reorder factor	levels of month	to be from Jan to Dec. Add one column to the data indicating	
whether	a	fire occurred	for each observation.	(True	for	area	> 0 and False	for	area == 0)

```{r q4-b}
ff.df$x <- ff.df$area > 0

set.seed(12)
df <- data.frame(month = sample(12), x = LETTERS[1:12])

library(dplyr)
ff.df %>% 
   mutate(month = factor(month.abb[month], levels = month.abb)) %>% 
   arrange(month)
```
