---
title: "Data Analysis - Lab-2"
author: "Doug Duril"
date: "June 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Create two matrices from the given set of numbers:
X1 = {2,3,7,1,6,2,3,5,1} and X2 = {3,2,9,0,7,8,5,8,2}

``` {r q1}
X1 <- matrix(c(2,3,7,1,6,2,3,5,1), nrow=3, ncol=3)
X2 <- matrix(c(3,2,9,0,7,8,5,8,2), nrow=3, ncol=3)
X1
X2
```

2. Find the matrix product.

``` {r q2}
X3 <- X1 %*% X2
X3
```

3. Find the inverse of the matrix and prove that it is correct.

``` {r q3}
solve(X3)

iX3 <- solve(X3)
X3 %*% iX3
```

4. Load the Animals data set from the "MASS" package.

``` {r q4}
library(MASS)
head(Animals)
```

5. Find the correlation coefficient of brain and body in this data set and comment on the relation between them.

``` {r q5}
body <- Animals$body
brain <- Animals$brain
cor(brain, body)

cor(Animals)
```

6. Load USArrests data set. 

Load USArrests and run a few functions to review dataset - head, names, list data.

``` {r q6-a}
data(USArrests)
head(USArrests)
names(USArrests)
USArrests
```

Comment on the distribution of the variables. Is there any relation between UrbanPop and three crimes?

``` {r q6-b}
summary(USArrests)
```

``` {r q6-c}
cor(USArrests)
```

Running plot for these combinations, Murder and Assault do not appear to have a relation to UrbanPop. The distribution of plot points are scattered to the point that they do not appear to correlate to UrbanPop.

``` {r q6-d}
with(USArrests, plot(UrbanPop, Murder))
with(USArrests, plot(UrbanPop, Assault))
```
 	 
Plotting Rape against UrbanPop - there are some outliers, but there does seem to be a relationship between Rape and UrbanPop. There is an indication of a correlation and more analysis could be performed.

``` {r q6-e}
with(USArrests, plot(UrbanPop, Rape))
```
 
7. Which states has most and least assault, murder, and rape arrests?

``` {r q7}
# most and least assault
x <- which(USArrests$Assault == max(USArrests$Assault))
rownames(USArrests)[x]
x <- which(USArrests$Assault == min(USArrests$Assault))
rownames(USArrests)[x]

# most and least murder
x <- which(USArrests$Murder == max(USArrests$Murder))
rownames(USArrests)[x]
x <- which(USArrests$Murder == min(USArrests$Murder))
rownames(USArrests)[x]

# most and least rape
x <- which(USArrests$Rape == max(USArrests$Rape))
rownames(USArrests)[x]
x <- which(USArrests$Rape == min(USArrests$Rape))
rownames(USArrests)[x]
```

8. List the states which have assault arrests more than median of the country.

``` {r q8}
assault.median = median(USArrests$Assault)
assault.median
subset(USArrests, Assault > assault.median, select= c(UrbanPop, Assault))
```

9. Which states are in the bottom 25% of murder?

``` {r q9}
bottomQuartileMurderRate <- quantile(USArrests$Murder)[2]
bottomQuartileMurderRate
subset(USArrests, Murder < bottomQuartileMurderRate, select= c(UrbanPop, Murder))
```

10. Which states are in the top 25% of the murder.

``` {r q10}
topQuartileMurderRate <- quantile(USArrests$Murder)[4]
topQuartileMurderRate
subset(USArrests, Murder > topQuartileMurderRate, select= c(UrbanPop, Murder))
```
