---
title: "Data Analysis - Lab-1"
author: "Doug Duril"
date: "June 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. What are the measures of central tendency of data? What is the order of these measures for
skewed data?

Arithmetic mean (average), median (middle value for a set of data) and the mode (most frequent value in the set of data).

The average is the most popular measure, but when the data is skewed, the median is preferred. 

2. How can you measure variation of data?

The range can be used - which is the difference between the lowest and highest values in the data set. Quartiles can also be used and these divide the data set into equal quarters.

3. Create a vector of positive odd integers less than 100

```{r q3-a}
v = seq(1, 100, by=2)
v
```

Although, for learning, it might be useful to use this format:

```{r q3-b}
v = seq(from=1, to=100, by=2)
v
```

4. Remove the values greater than 60 and less than 80

```{r q4}
v = seq(from=1, to=100, by=2)
v <- v[v < 60 | v > 80]
v
```

5. Find the five number summary of the remaining set of values

```{r q5-a}
v = seq(from=1, to=100, by=2)
v <- v[v < 60 | v > 80]
fivenum(v)
```

Example where values 61 through 79 are omitted and fivenum() used to verify output is the same as above.

```{r q5-b}
v = c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29,31,33,35,37,39,41,43,45,47,49,51,53,55,57,59,81,83,85,87,89,91,93,95,97,99)
fivenum(v)
```

6. Consider the following vector of values. X <-{8, 14, 9, 15, NA, 8,13, 2,9, NA}
Write an R program to return the positions of the missing values

``` {r q6}
x <- c(8, 14, 9, 15, NA, 8, 13, 2, 9, NA)
which(is.na(x))
```

7. Write an R program to count the number of non-missing values

``` {r q7}
sum(x, na.rm = TRUE)
```

8. Write an R program to replace the missing values with the mean of the non-missing values.

``` {r q8}
x <- c(8, 14, 9, 15, NA, 8, 13, 2, 9, NA)
m <- mean(x, na.rm = TRUE)
x[is.na(x)] <- m
x
```

9. Load mtcars data. Write an R program that will rearrange the rows of the data frame so that they are sorted by the value of Horsepower.

``` {r q9}
df.hp <- mtcars[order(mtcars$hp),]
df.hp
```

10. Write an R program to count the number of observations with cylinders greater than 4 and gear greater than 3.

``` {r q10}
mtcars

df.x = which(mtcars$cyl > 4 & mtcars$gear > 3)
df.x

x <- length(which(mtcars$cyl > 4 & mtcars$gear > 3))
x
```
